---
title: "Untitled"
author: "Di"
date: "2025-03-26"
output: 
  html_document:
    df_print: paged
    code_folding: show
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: true
    theme: simplex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```

# Introduction

This report provides a statistical analysis of employee data to understand key factors influencing attrition, performance, and satisfaction. We leverage two primary datasets: one detailing employee demographic and job information, and another containing performance review data over time. The analysis follows a structured approach, focusing on statistical testing and modeling to derive actionable insights.

```{r}
# Load libraries
library(readr)
library(dplyr)
library(lubridate) # For handling dates
library(tidyr)     # For data manipulation like pivot_wider
library(broom)     # For tidying model outputs
library(car)
library(MASS)
library(lmtest)
```



## Data Loading

```{r}
employee_data <- read.csv("Employee.csv")
performance_rating_data <- read_csv("PerformanceRating.csv", show_col_types = FALSE)

rating_levels <- read.csv("RatingLevel.csv")
satisfaction_levels <- read.csv("SatisfiedLevel.csv")
education_levels <- read.csv("EducationLevel.csv")
```


## Data Cleaning and Merging

```{r}

# Convert date columns 
employee_data <- employee_data %>%
  mutate(HireDate = as.Date(HireDate, format = "%m/%d/%Y"))

performance_rating_data <- performance_rating_data %>%
  mutate(ReviewDate = as.Date(ReviewDate, format = "%m/%d/%Y"))

# Get the most recent review per employee
latest_performance <- performance_rating_data %>%
  arrange(EmployeeID, desc(ReviewDate)) %>%
  distinct(EmployeeID, .keep_all = TRUE)

# Merge data. This keeps all 1470 employees; those without performance data will have NA for performance fields
combined_data <- employee_data %>%
  left_join(latest_performance, by = "EmployeeID") %>%
  # Join lookup tables using the correct ID columns
  left_join(education_levels, by = c("Education" = "EducationLevelID")) %>%
  left_join(satisfaction_levels, by = c("EnvironmentSatisfaction" = "SatisfactionID")) %>%
  rename(EnvironmentSatisfactionLevel = SatisfactionLevel) %>%
  left_join(satisfaction_levels, by = c("JobSatisfaction" = "SatisfactionID")) %>%
  rename(JobSatisfactionLevel = SatisfactionLevel) %>%
  left_join(satisfaction_levels, by = c("RelationshipSatisfaction" = "SatisfactionID")) %>%
  rename(RelationshipSatisfactionLevel = SatisfactionLevel) %>%
  left_join(rating_levels, by = c("SelfRating" = "RatingID")) %>%
  rename(SelfRatingLevel = RatingLevel) %>%
  left_join(rating_levels, by = c("ManagerRating" = "RatingID")) %>%
  rename(ManagerRatingLevel = RatingLevel)

# Convert relevant columns to factors for modeling
combined_data <- combined_data %>%
  mutate(
    Gender = factor(Gender),
    BusinessTravel = factor(BusinessTravel),
    Department = factor(Department),
    Ethnicity = factor(Ethnicity),
    EducationField = factor(EducationField),
    JobRole = factor(JobRole),
    MaritalStatus = factor(MaritalStatus),
    OverTime = factor(OverTime),
    # Ensure 'Yes' is the second level for logistic regression interpretation (odds of 'Yes')
    Attrition = factor(Attrition, levels = c("No", "Yes")),
    # Use ordered factors for variables with inherent order
    EducationLevel = factor(EducationLevel, levels = c("No Formal Qualifications", "High School", "Bachelors", "Masters", "Doctorate"), ordered = TRUE),
    StockOptionLevel = factor(StockOptionLevel, ordered = TRUE),
    EnvironmentSatisfaction = factor(EnvironmentSatisfaction, ordered = TRUE),
    JobSatisfaction = factor(JobSatisfaction, ordered = TRUE),
    RelationshipSatisfaction = factor(RelationshipSatisfaction, ordered = TRUE),
    WorkLifeBalance = factor(WorkLifeBalance, ordered = TRUE),
    SelfRating = factor(SelfRating, ordered = TRUE),
    ManagerRating = factor(ManagerRating, ordered = TRUE)
    )

# Create numeric versions for analyses treating them as continuous/pseudo-continuous
combined_data <- combined_data %>%
  mutate(
    ManagerRatingNumeric = as.numeric(ManagerRating),
    JobSatisfactionNumeric = as.numeric(JobSatisfaction),
    EnvironmentSatisfactionNumeric = as.numeric(EnvironmentSatisfaction),
    WorkLifeBalanceNumeric = as.numeric(WorkLifeBalance)
  )


# Summary of the final dataset
unique_employees_in_performance <- n_distinct(performance_rating_data$EmployeeID)
total_employees <- nrow(employee_data)
employees_without_performance <- total_employees - unique_employees_in_performance

print(paste("Number of unique EmployeeIDs in performance_data:", unique_employees_in_performance))
print(paste("Number of rows (employees) in employee_data:", total_employees))
print(paste("Number of employees in employee_data WITHOUT records in performance_data:", employees_without_performance))

```


# Attrition Analysis

## Statistical Tests (Group Comparisons)

We use t-tests (for numeric variables, assuming approximate normality or large sample size) or Wilcoxon tests (non-parametric alternative), and Chi-squared tests (for categorical variables) to compare employees who attrited versus those who did not.

```{r}
# Define fixed variables for demonstration instead of using Shiny inputs
num_var_attr_fixed <- "Salary"
cat_var_attr_fixed <- "OverTime"

# --- Numeric/Ordinal Test ---
cat("\n--- Test for Numeric Variable:", num_var_attr_fixed, "---\n")
if (!is.null(num_var_attr_fixed) && num_var_attr_fixed %in% names(combined_data) && is.numeric(combined_data[[num_var_attr_fixed]])) {
    formula_num <- as.formula(paste(num_var_attr_fixed, "~ Attrition"))

    # Perform t-test (Welch's)
    t_test_result <- tryCatch({
        t.test(formula_num, data = combined_data, var.equal = FALSE)
    }, error = function(e) NULL)

    if (!is.null(t_test_result)) {
        cat(paste("--- Welch's T-Test for", num_var_attr_fixed, "by Attrition ---\n"))
        print(tidy(t_test_result))

        # Calculate and print means
        means <- combined_data %>%
            filter(!is.na(Attrition) & !is.na(.data[[num_var_attr_fixed]])) %>%
            group_by(Attrition) %>%
            summarise(Mean = mean(.data[[num_var_attr_fixed]], na.rm = TRUE), N = n(), .groups = 'drop')
        cat("\nGroup Means:\n")
        print(means)

    } else {
        cat(paste("Could not perform t-test for", num_var_attr_fixed, ". Attempting Wilcoxon test...\n"))
        wilcox_result <- tryCatch({
            wilcox.test(formula_num, data = combined_data)
        }, error = function(e) NULL)
        if (!is.null(wilcox_result)){
            cat(paste("\n--- Wilcoxon Rank Sum Test for", num_var_attr_fixed, "by Attrition ---\n"))
            print(tidy(wilcox_result))
        } else {
            cat(paste("Could not perform Wilcoxon test for", num_var_attr_fixed, "either.\n"))
        }
    }
    cat("\n") # Add spacing
} else {
    cat(paste("Selected numeric variable", num_var_attr_fixed, "not found or not suitable for testing.\n\n"))
}

# --- Categorical Test ---
cat("\n--- Test for Categorical Variable:", cat_var_attr_fixed, "---\n")
if (!is.null(cat_var_attr_fixed) && cat_var_attr_fixed %in% names(combined_data) && (is.factor(combined_data[[cat_var_attr_fixed]]) || is.character(combined_data[[cat_var_attr_fixed]]))) {
    cat(paste("\n--- Chi-Squared Test for", cat_var_attr_fixed, "by Attrition ---\n"))
    # Create contingency table
    cont_table <- tryCatch({
        table(combined_data[[cat_var_attr_fixed]], combined_data$Attrition, useNA = "no")
    }, error = function(e) NULL)

    if (is.null(cont_table) || nrow(cont_table) < 2 || ncol(cont_table) < 2) {
        cat("Could not create valid contingency table.\n")
    } else {
        # Perform Chi-squared test
        chisq_result <- tryCatch({
            chisq.test(cont_table, simulate.p.value = TRUE) # Simulate p-value for robustness
        }, error = function(e) NULL)

        if (!is.null(chisq_result)) {
            cat("Contingency Table (excluding NAs):\n")
            print(cont_table)
            cat("\nTest Result (with simulated p-value if needed):\n")
            print(tidy(chisq_result))
            # Check expected counts warning
            if(any(chisq_result$expected < 5)) {
                cat("\nWarning: Expected counts < 5 exist, Chi-squared approximation may be inaccurate (simulation used).\n")
            }
        } else {
            cat(paste("Could not perform Chi-Squared test for", cat_var_attr_fixed, ".\n"))
        }
    }
 } else {
     cat(paste("\nSelected categorical variable", cat_var_attr_fixed, "not found or not a factor/character.\n"))
 }
```

## Logistic Regression Model

We build a logistic regression model to predict the probability of attrition based on multiple factors simultaneously.


```{r attrition-logistic-regression-corrected-masking}
# Define fixed predictors for demonstration
selected_preds_attr_fixed <- c("Age", "OverTime", "JobSatisfaction", "Salary", "YearsAtCompany", "MaritalStatus", "DistanceFromHome..KM.", "EnvironmentSatisfaction")

# Ensure these predictors exist in the data
valid_predictors_attr_fixed <- intersect(selected_preds_attr_fixed, names(combined_data))

if (length(valid_predictors_attr_fixed) == 0) {
    cat("None of the selected fixed predictors exist in the data.")
} else {
    # Create the formula string dynamically
    formula_str_attr <- paste("Attrition ~", paste(valid_predictors_attr_fixed, collapse = " + "))
    formula_obj_attr <- as.formula(formula_str_attr)

    cat("--- Logistic Regression Model for Attrition ---\n")
    cat("Formula:", formula_str_attr, "\n\n")

    # CORRECTED AGAIN: Explicitly use dplyr::select to avoid masking
    model_data_attr <- combined_data %>%
        # Use dplyr::select explicitly
        dplyr::select(any_of(c("Attrition", valid_predictors_attr_fixed))) %>%
        na.omit() # Remove rows with NA in *any* selected column

    if(nrow(model_data_attr) == 0) {
        cat("No complete cases found for the selected predictors and outcome. Cannot fit model.\n")
    } else {
        cat(paste("Fitting model using", nrow(model_data_attr), "complete observations.\n\n"))

        # Fit the model
        logistic_model_attr <- tryCatch({
            # Use the filtered model_data_attr which only contains necessary columns
            glm(formula_obj_attr, data = model_data_attr, family = binomial(link = "logit"))
        }, error = function(e) {
            cat("Error fitting model:", conditionMessage(e), "\n")
            NULL
        })

        if (!is.null(logistic_model_attr)) {
            # Print model summary
            cat("Model Summary:\n")
            print(summary(logistic_model_attr))

            # Print tidy coefficients with Odds Ratios
            cat("\nCoefficients (Tidy) with Odds Ratios:\n")
            tidy_coeffs <- tidy(logistic_model_attr, conf.int = TRUE, exponentiate = TRUE)
            print(tidy_coeffs)

            cat("\nInterpretation of Odds Ratios (exp(estimate)):\n")
            cat("- OR > 1: Odds of Attrition ('Yes') increase.\n")
            cat("- OR < 1: Odds of Attrition ('Yes') decrease.\n")
            cat("- Confidence interval includes 1: Not statistically significant (p > 0.05).\n")

            cat("\nAIC:", AIC(logistic_model_attr), "\n")

        } else {
            cat("Model could not be fitted with the selected predictors.\n")
        }
    }
}
```

# Performance Analysis

## Correlation Analysis

Examine correlations between satisfaction metrics, work-life balance, training, and manager ratings.

```{r performance-correlation-spearman-corrected-masking}
# Data preparation remains the same, selecting numeric/pseudo-numeric versions
corr_data_perf <- combined_data %>%
  mutate(
     SelfRatingNumeric = as.numeric(SelfRating) # Convert for correlation if desired
     ) %>%
  # Explicitly use dplyr::select
  dplyr::select(any_of(c("ManagerRatingNumeric", "SelfRatingNumeric", "JobSatisfactionNumeric",
                  "EnvironmentSatisfactionNumeric", "RelationshipSatisfactionNumeric",
                  "WorkLifeBalanceNumeric", "TrainingOpportunitiesTaken", "Salary",
                  "YearsAtCompany", "YearsInMostRecentRole", "Age", "DistanceFromHome..KM."))) %>%
  # Keep select_if for now, though superseded by `where()` or `across()`
  select_if(is.numeric)

valid_corr_vars <- names(corr_data_perf)

# Define fixed variables for correlation matrix
selected_vars_corr_fixed <- intersect(c("ManagerRatingNumeric", "JobSatisfactionNumeric", "EnvironmentSatisfactionNumeric", "WorkLifeBalanceNumeric", "YearsAtCompany", "Salary"), valid_corr_vars)

if (length(selected_vars_corr_fixed) < 2) {
    cat("Need at least two valid variables for correlation.")
} else {
    # Ensure selected variables exist in the numeric dataset
    # Base R subsetting is not affected by masking
    selected_data_corr <- corr_data_perf[, selected_vars_corr_fixed, drop = FALSE]

    # Calculate Spearman's rank correlation matrix
    # Handles missing data pair-wise and uses ranks (suitable for ordinal)
    cor_matrix <- cor(selected_data_corr, use = "pairwise.complete.obs", method = "spearman")

    cat("--- Spearman's Rank Correlation Matrix (rho) ---\n")
    cat("Variables:", paste(selected_vars_corr_fixed, collapse=", "), "\n\n")
    print(round(cor_matrix, 3))

    cat("\nInterpretation: Spearman's rho measures the strength and direction of monotonic association between variables.\n")
    cat("Values close to 1 indicate strong positive monotonic relationship, close to -1 strong negative monotonic relationship, close to 0 weak monotonic relationship.\n")
}
```

## Ordinal Logistic Regression Model for Manager Rating

Build an ordinal logistic regression model to predict the ordered ManagerRating based on potential influencing factors. This is more appropriate than linear regression for an ordinal outcome.

```{r performance-ordinal-regression-corrected-masking}
# Ensure the outcome variable is the original ordered factor
outcome_var_perf_ord <- "ManagerRating"

# Define fixed predictors (use original factors where appropriate)
selected_preds_perf_ord_fixed <- intersect(c("JobSatisfaction", "EnvironmentSatisfaction", "WorkLifeBalance", "YearsAtCompany", "SelfRating", "TrainingOpportunitiesTaken"), names(combined_data))

if (length(selected_preds_perf_ord_fixed) == 0) {
    cat("No valid fixed predictors found for the Manager Rating model.")
# Ensure the outcome variable is an ordered factor
} else if (!outcome_var_perf_ord %in% names(combined_data) || !is.ordered(combined_data[[outcome_var_perf_ord]])) {
    cat("Outcome variable '", outcome_var_perf_ord, "' not found or is not an ordered factor.")
} else {
    # Create formula using the original ORDERED factor as the outcome
    formula_str_perf_ord <- paste(outcome_var_perf_ord, "~", paste(selected_preds_perf_ord_fixed, collapse = " + "))
    formula_obj_perf_ord <- as.formula(formula_str_perf_ord)

    cat("--- Ordinal Logistic Regression Model for Manager Rating ---\n")
    cat("Formula:", formula_str_perf_ord, "\n\n")

    # CORRECTED AGAIN: Filter data using dplyr::select explicitly
    model_data_perf_ord <- combined_data %>%
        # Explicitly call dplyr::select
        dplyr::select(all_of(c(outcome_var_perf_ord, selected_preds_perf_ord_fixed))) %>%
        na.omit() # Remove rows with NA in any selected column

    # Ensure sufficient data and levels in the outcome
    if(nrow(model_data_perf_ord) == 0) {
        cat("No complete cases found for the selected variables. Cannot fit model.\n")
    } else if (nlevels(model_data_perf_ord[[outcome_var_perf_ord]]) < 2) {
        cat("Outcome variable has fewer than 2 levels after removing NAs. Cannot fit model.\n")
    } else {
        cat(paste("Fitting model using", nrow(model_data_perf_ord), "complete observations.\n\n"))

        # Fit the Ordinal Logistic Regression model using polr from MASS package
        # Hess=TRUE is needed to calculate standard errors for summary/coeftest
        polr_model_perf <- tryCatch({
            polr(formula_obj_perf_ord, data = model_data_perf_ord, Hess = TRUE)
        }, warning = function(w){ # Catch warnings e.g. about design matrix
            cat("Warning during model fitting:", conditionMessage(w), "\n")
            # Attempt to fit anyway
            suppressWarnings(polr(formula_obj_perf_ord, data = model_data_perf_ord, Hess = TRUE))
        }, error = function(e) {
            cat("Error fitting Ordinal Logistic Regression model:", conditionMessage(e), "\n")
            NULL
        })

        if (!is.null(polr_model_perf)) {
            cat("Model Summary (Coefficients are log-odds):\n")
            # summary() gives coefficients (log-odds) and intercepts (cut-points)
            print(summary(polr_model_perf))

            cat("\nCoefficients with Standard Errors and P-values (from coeftest):\n")
            # Use coeftest from lmtest package for p-values
            coeffs_test <- tryCatch(coeftest(polr_model_perf), error = function(e) NULL)
            if (!is.null(coeffs_test)) {
               print(coeffs_test)
            } else {
               cat("Could not calculate p-values using coeftest.\n")
            }


            cat("\nOdds Ratios (Exponentiated Coefficients):\n")
            # Use tidy with exponentiate = TRUE for Odds Ratios
            tidy_coeffs_or <- tidy(polr_model_perf, exponentiate = TRUE, conf.int = TRUE)
            print(tidy_coeffs_or)

            cat("\nInterpretation of Odds Ratios (OR):\n")
            cat("- OR > 1: For a one-unit increase in the predictor (or for the specific factor level vs reference), the odds of being in a *higher* Manager Rating category (vs. all lower categories combined) increase by a factor of OR, holding other predictors constant.\n")
            cat("- OR < 1: The odds of being in a *higher* Manager Rating category decrease.\n")
            cat("- Confidence interval includes 1: Predictor is likely not statistically significant (check p-value from coeftest).\n")
            cat("- Intercepts (zeta coefficients) define the boundaries between rating categories on the log-odds scale.\n")
            cat("\nAIC:", AIC(polr_model_perf), "\n")

        } else {
            cat("Ordinal Logistic Regression model could not be fitted.\n")
        }
    }
}
```

## Comparing Ratings/Satisfaction Across Groups (ANOVA)

Use Analysis of Variance (ANOVA) to test if the average manager rating (or satisfaction scores) differs significantly across different groups (e.g., Departments, Job Roles).

```{r}
# Define fixed variables for ANOVA
outcome_anova_perf_fixed <- "ManagerRatingNumeric"
group_anova_perf_fixed <- "Department"

# Checks
if (!outcome_anova_perf_fixed %in% names(combined_data) || !group_anova_perf_fixed %in% names(combined_data)) {
    cat("Selected variable(s) for ANOVA not found in the data.\n")
} else if (!is.numeric(combined_data[[outcome_anova_perf_fixed]])) {
    cat(paste("Outcome variable", outcome_anova_perf_fixed, "is not numeric.\n"))
} else if (!is.factor(combined_data[[group_anova_perf_fixed]]) && !is.character(combined_data[[group_anova_perf_fixed]])) {
    cat(paste("Grouping variable", group_anova_perf_fixed, "is not categorical.\n"))
} else {
    formula_anova_perf <- as.formula(paste(outcome_anova_perf_fixed, "~", group_anova_perf_fixed))

    cat(paste("--- ANOVA for", outcome_anova_perf_fixed, "by", group_anova_perf_fixed, "---\n"))
    cat("Formula:", deparse(formula_anova_perf), "\n\n")

    # Filter data
    anova_data_perf <- combined_data %>%
        select(all_of(c(outcome_anova_perf_fixed, group_anova_perf_fixed))) %>%
        na.omit()

    # Check data sufficiency
    if (length(unique(anova_data_perf[[group_anova_perf_fixed]])) < 2) {
        cat("Grouping variable has fewer than 2 levels with data. Cannot perform ANOVA.\n")
    } else if (nrow(anova_data_perf) == 0) {
        cat("No complete observations for this combination of variables.\n")
    } else {
        min_group_size <- tryCatch(min(table(anova_data_perf[[group_anova_perf_fixed]])), error=function(e) 0)
        if (min_group_size < 2) {
             cat("Warning: At least one group has fewer than 2 observations. ANOVA results may be unreliable.\n")
        }
        cat(paste("Performing ANOVA using", nrow(anova_data_perf), "complete observations.\n\n"))

        # Perform ANOVA using lm and car::Anova
        anova_model_perf <- tryCatch({
            lm_for_anova <- lm(formula_anova_perf, data = anova_data_perf)
            Anova(lm_for_anova, type = "II")
        }, error = function(e) {
            cat("Error fitting ANOVA model:", conditionMessage(e), "\n")
            NULL
        })

        if (!is.null(anova_model_perf)) {
            cat("ANOVA Summary (Type II SS):\n")
            print(anova_model_perf)

            p_value <- tryCatch(anova_model_perf[group_anova_perf_fixed, "Pr(>F)"], error = function(e) NA)

            if (!is.na(p_value) && p_value < 0.05) {
                cat(paste("\nSignificant difference found across groups (p =", format.pval(p_value, digits = 3), ").\n"))

                # Perform post-hoc test if significant AND more than 2 groups
                if (length(unique(anova_data_perf[[group_anova_perf_fixed]])) > 2) {
                    cat("Performing post-hoc tests (Tukey HSD)...\n")
                    aov_model_for_tukey <- tryCatch(aov(formula_anova_perf, data = anova_data_perf), error = function(e) NULL)
                    if (!is.null(aov_model_for_tukey)) {
                        tukey_result <- tryCatch({ TukeyHSD(aov_model_for_tukey, which = group_anova_perf_fixed) }, error = function(e) NULL)

                        if(!is.null(tukey_result)){
                            tidy_tukey <- tidy(tukey_result)
                            significant_tukey <- tidy_tukey %>% filter(adj.p.value < 0.05)
                            if (nrow(significant_tukey) > 0) {
                                print(significant_tukey)
                                cat("\nAbove table shows specific group pairs that differ significantly (adj.p.value < 0.05).\n")
                            } else {
                                cat("\nNo specific group pairs found to be significantly different after adjustment.\n")
                            }
                        } else {
                            cat("Could not perform Tukey HSD post-hoc test.\n")
                        }
                    } else {
                        cat("Could not create model required for Tukey HSD.\n")
                    }
                }
            } else if (!is.na(p_value)) {
                cat(paste("\nNo significant difference found across groups (p =", format.pval(p_value, digits = 3), ").\n"))
            } else {
                cat("\nCould not determine significance (p-value missing or error).\n")
            }
        } else {
            cat("ANOVA model could not be fitted.\n")
        }
    }
}
```

# Satisfaction Analysis

## Comparing Satisfaction Across Groups (ANOVA)

Use Analysis of Variance (ANOVA) to test if average satisfaction scores differ significantly across different categorical groups (e.g., Department, Job Role, OverTime status).

```{r}
# Define fixed variables for Satisfaction ANOVA
outcome_anova_satisfaction_fixed <- "JobSatisfactionNumeric"
group_anova_satisfaction_fixed <- "JobRole"

# Checks
if (!outcome_anova_satisfaction_fixed %in% names(combined_data) || !group_anova_satisfaction_fixed %in% names(combined_data)) {
    cat("Selected variable(s) for Satisfaction ANOVA not found.\n")
} else if (!is.numeric(combined_data[[outcome_anova_satisfaction_fixed]])) {
    cat(paste("Outcome variable", outcome_anova_satisfaction_fixed, "is not numeric.\n"))
} else if (!is.factor(combined_data[[group_anova_satisfaction_fixed]]) && !is.character(combined_data[[group_anova_satisfaction_fixed]])) {
    cat(paste("Grouping variable", group_anova_satisfaction_fixed, "is not categorical.\n"))
} else {

    formula_anova_satisfaction <- as.formula(paste(outcome_anova_satisfaction_fixed, "~", group_anova_satisfaction_fixed))

    cat(paste("--- ANOVA for", outcome_anova_satisfaction_fixed, "by", group_anova_satisfaction_fixed, "---\n"))
    cat("Formula:", deparse(formula_anova_satisfaction), "\n\n")

    # Filter data
    anova_data_satisfaction <- combined_data %>%
        select(all_of(c(outcome_anova_satisfaction_fixed, group_anova_satisfaction_fixed))) %>%
        na.omit()

    # Check data sufficiency
    if (length(unique(anova_data_satisfaction[[group_anova_satisfaction_fixed]])) < 2) {
        cat("Grouping variable has fewer than 2 levels with data. Cannot perform ANOVA.\n")
    } else if (nrow(anova_data_satisfaction) == 0) {
        cat("No complete observations for this combination.\n")
    } else {
        min_group_size <- tryCatch(min(table(anova_data_satisfaction[[group_anova_satisfaction_fixed]])), error=function(e) 0)
        if (min_group_size < 2) {
             cat("Warning: At least one group has fewer than 2 observations.\n")
        }
        cat(paste("Performing ANOVA using", nrow(anova_data_satisfaction), "complete observations.\n\n"))

        # Perform ANOVA
        anova_model_satisfaction <- tryCatch({
            lm_for_anova_sat <- lm(formula_anova_satisfaction, data = anova_data_satisfaction)
            Anova(lm_for_anova_sat, type = "II")
        }, error = function(e) {
            cat("Error fitting ANOVA model:", conditionMessage(e), "\n")
            NULL
        })

        if (!is.null(anova_model_satisfaction)) {
            cat("ANOVA Summary (Type II SS):\n")
            print(anova_model_satisfaction)

            p_value_sat <- tryCatch(anova_model_satisfaction[group_anova_satisfaction_fixed,"Pr(>F)"], error=function(e) NA)

            if (!is.na(p_value_sat) && p_value_sat < 0.05) {
                cat(paste("\nSignificant difference in", outcome_anova_satisfaction_fixed, "found across groups (p =", format.pval(p_value_sat, digits=3), ").\n"))

                if (length(unique(anova_data_satisfaction[[group_anova_satisfaction_fixed]])) > 2) {
                    cat("Performing post-hoc tests (Tukey HSD)...\n")
                    aov_model_for_tukey_sat <- tryCatch(aov(formula_anova_satisfaction, data = anova_data_satisfaction), error = function(e) NULL)
                    if (!is.null(aov_model_for_tukey_sat)) {
                        tukey_result_sat <- tryCatch({ TukeyHSD(aov_model_for_tukey_sat, which = group_anova_satisfaction_fixed) }, error = function(e) NULL)

                        if(!is.null(tukey_result_sat)){
                            tidy_tukey_sat <- tidy(tukey_result_sat)
                            significant_tukey_sat <- tidy_tukey_sat %>% filter(adj.p.value < 0.05)
                            if (nrow(significant_tukey_sat) > 0) {
                                print(significant_tukey_sat)
                                cat("\nAbove table shows specific group pairs that differ significantly in", outcome_anova_satisfaction_fixed, "(adj.p.value < 0.05).\n")
                            } else {
                                cat("\nNo specific group pairs found to be significantly different after adjustment.\n")
                            }
                        } else {
                            cat("Could not perform Tukey HSD post-hoc test.\n")
                        }
                    } else {
                        cat("Could not create model required for Tukey HSD.\n")
                    }
                }
            } else if (!is.na(p_value_sat)) {
                cat(paste("\nNo significant difference in", outcome_anova_satisfaction_fixed, "found across groups (p =", format.pval(p_value_sat, digits=3), ").\n"))
            } else {
                 cat("\nCould not determine significance (p-value missing or error).\n")
            }

        } else {
            cat("ANOVA model could not be fitted.\n")
        }
    }
}
```


## Ordinal Logistic Regression Model for Satisfaction

Model a selected satisfaction metric (which is ordinal) based on multiple predictors using Ordinal Logistic Regression. 

```{r satisfaction-ordinal-regression}
# Define fixed outcome (use the original ORDERED factor)
selected_outcome_sat_ord_fixed <- "JobSatisfaction" # e.g., JobSatisfaction, EnvironmentSatisfaction, RelationshipSatisfaction

# Define fixed predictors (Using ManagerRating factor instead of numeric)
selected_preds_sat_ord_fixed <- intersect(c("Age", "OverTime", "JobRole", "YearsWithCurrManager", "WorkLifeBalance", "Salary", "ManagerRating"), names(combined_data)) # Use ManagerRating factor

# Checks
if (length(selected_preds_sat_ord_fixed) == 0) {
    cat("No valid fixed predictors found for the Satisfaction model.")
# Check if outcome exists and is an ordered factor
} else if (!selected_outcome_sat_ord_fixed %in% names(combined_data) || !is.ordered(combined_data[[selected_outcome_sat_ord_fixed]])) {
    cat(paste("Outcome variable '", selected_outcome_sat_ord_fixed, "' not found or is not an ordered factor.\n"))
} else {
    # Create formula using the original ORDERED factor
    formula_str_satisfaction_ord <- paste(selected_outcome_sat_ord_fixed, "~", paste(selected_preds_sat_ord_fixed, collapse = " + "))
    formula_obj_satisfaction_ord <- as.formula(formula_str_satisfaction_ord)

    cat(paste("--- Ordinal Logistic Regression Model for", selected_outcome_sat_ord_fixed, "---\n"))
    cat("Formula:", formula_str_satisfaction_ord, "\n\n")

    # Filter data for complete cases
    model_data_satisfaction_ord <- combined_data %>%
        select(all_of(c(selected_outcome_sat_ord_fixed, selected_preds_sat_ord_fixed))) %>%
        na.omit()

    # Check data sufficiency
    if(nrow(model_data_satisfaction_ord) == 0) {
        cat("No complete cases found for the selected variables. Cannot fit model.\n")
     } else if (nlevels(model_data_satisfaction_ord[[selected_outcome_sat_ord_fixed]]) < 2) {
        cat("Outcome variable has fewer than 2 levels after removing NAs. Cannot fit model.\n")
    } else {
        cat(paste("Fitting model using", nrow(model_data_satisfaction_ord), "complete observations.\n\n"))

        # Fit the Ordinal Logistic Regression model
        polr_model_satisfaction <- tryCatch({
            polr(formula_obj_satisfaction_ord, data = model_data_satisfaction_ord, Hess = TRUE)
         }, warning = function(w){
            cat("Warning during model fitting:", conditionMessage(w), "\n")
            suppressWarnings(polr(formula_obj_satisfaction_ord, data = model_data_satisfaction_ord, Hess = TRUE))
        }, error = function(e) {
            cat("Error fitting Ordinal Logistic Regression model:", conditionMessage(e), "\n")
            NULL
        })

        if (!is.null(polr_model_satisfaction)) {
            cat("Model Summary (Coefficients are log-odds):\n")
            print(summary(polr_model_satisfaction))

            cat("\nCoefficients with Standard Errors and P-values (from coeftest):\n")
            coeffs_test_sat <- tryCatch(coeftest(polr_model_satisfaction), error = function(e) NULL)
             if (!is.null(coeffs_test_sat)) {
               print(coeffs_test_sat)
            } else {
               cat("Could not calculate p-values using coeftest.\n")
            }

            cat("\nOdds Ratios (Exponentiated Coefficients):\n")
            tidy_coeffs_or_sat <- tidy(polr_model_satisfaction, exponentiate = TRUE, conf.int = TRUE)
            print(tidy_coeffs_or_sat)

            cat("\nInterpretation of Odds Ratios (OR):\n")
            cat("- OR > 1: Odds of being in a *higher* satisfaction category increase.\n")
            cat("- OR < 1: Odds of being in a *higher* satisfaction category decrease.\n")
            cat("- Confidence interval includes 1: Predictor likely not statistically significant (check p-value).\n")
            cat("- Intercepts define boundaries between satisfaction categories (log-odds scale).\n")
            cat("\nAIC:", AIC(polr_model_satisfaction), "\n")

        } else {
            cat("Ordinal Logistic Regression model could not be fitted.\n")
        }
    }
}